## Exercise 2.2
Heart Transplants

proportion of patients in the treatment group die

```{r}
30/34
45/69
```

1. What are the claims being tested? Use the same null and alternative hypothesis notation used in the section.

The null hypothesis, H_0, is that p_control = p_treatment. The alternative hypothesis, H_alpha, is that p_control > p_treatment. p is the proportion of 
dead people.

2. We write alive on `28` cards representing patients who were alive at the end of the study, and dead on `75` cards representing patients 
who were not. Then, we shuffle these cards and split them into two groups: one group of size `69` representing treatment, 
and another group of size `34` representing control. We calculate the difference between the proportion of dead cards in the 
treatment and control groups (treatment - control) and record this value. We repeat this many times to build a distribution 
centered at `null hypothesis`. Lastly, we calculate the fraction of simulations where the simulated differences in proportions are `p_c - p-t > 0`. If this 
fraction is low, we conclude that it is unlikely to have observed such an outcome by chance and that the null hypothesis should be 
rejected in favor of the alternative.

3. 41/99

## Exercise 2.6

1. The null hypothesis, H_0, is that p_c = p_t. The alternative hypothesis, H_alpha, is that p_control < p_treatment. 
p is the proportion of people who yawn.

2. 0.04411765 
    p_t = 10/34
    p_c = 4/16
    p_t - p_c 
```{r}
(10/34) - (4/16)
```

3. 0.4955/1.0. Since the p-value is higher than 0.05. We can't reject the null hypothesis and so, the control group doesn't have any effect.

