

```{r}
data <- read.csv("C:/Users/ds3/Desktop/yhm-coursework/week2/magnets.csv")

head(data)

summary(data)
```


1. What is the sample average of the change in score between the patients rating before the application of the device and the rating after the application?

> The sample average is 3.5.


2. Is the variable active a factor or a numeric variable?

> The variable active is a factor.


3. Compute the average value of the variable change for the patients that
 received and active magnet and average value for those that received an
 inactive placebo. (Hint: Notice that the rst 29 patients received an active
 magnet and the last 21 patients received an inactive placebo. The sub
sequence of the rst 29 values of the given variables can be obtained via
 the expression change[1:29] and the last 21 vales are obtained via the
 expression change[30:50] .)

 ```{r}
mean(data[,"change"][1:29])
mean(data[, "change"][30:50])
```

> The average value of the variable change for the patients that received 
        and active magnet is 5.241379. The average value for those that received
        an inactive placebo is 1.095238.




4. Compute the sample standard deviation of the variable change for the
patients that received and active magnet and the sample standard devia
tion for those that received an inactive placebo.

```{r}
sd(data[,"change"][1:29])
sd(data[, "change"][30:50])
```

> Received and active magnet: 3.236568. Received and inactive magnet: 1.578124.

5. Produce a boxplot of the variable change for the patients that received
 and active magnet and for patients that received an inactive placebo.
 What is the number of outliers in each subsequence?

```{r}
boxplot(data[,"change"][1:29])
boxplot(data[, "change"][30:50])
table(data[, "change"][30:50])
```

The number of outliers in patients that received an inactive placebo is 4.



Question 10.1. In Subsection 10.3.2 we compare the average against the mid
range as estimators of the expectation of the measurement. The goal of this
 exercise is to repeat the analysis, but this time compare the average to the
 median as estimators of the expectation in symmetric distributions.


 1. Simulate the sampling distribution of average and the median of a sample
 of size n = 100 from the Normal(3,2) distribution. Compute the expec
tation and the variance of the sample average and of the sample median.
 Which of the two estimators has a smaller mean square error

 ```{r}
library(tidyverse)

X_mean <- rep(0, 10^5)
X_medium <- rep(0, 10^5)

for (i in 1 : 10^5) {
        X <- rnorm(100, 3, sqrt(2))
        X_mean[i] <- mean(X)
        X_medium[i] <- median(X)
}

mean(X_mean)
mean(X_medium)

var(X_mean)
var(X_medium)
```

The sampling distribution of average of a sample has smaller mean square error than the sampling distribution of median of a sample.

2. Simulate the sampling distribution of average and the median of a sample
 of size n = 100 from the Uniform(0.5,5.5) distribution. Compute the
 expectation and the variance of the sample average and of the sample
 median. Which of the two estimators has a smaller mean square error?


```{r}
X_mean_unif <- rep(0, 10^5)
X_medium_unif <- rep(0, 10^5)

for (i in 1 : 10^5) {
        X <- runif(100, 0.5, sqrt(5.5))
        X_mean_unif[i] <- mean(X)
        X_medium_unif[i] <- median(X)
}

mean(X_mean_unif)
mean(X_medium_unif)

var(X_mean_unif)
var(X_medium_unif)

```


The sampling distribution of average of a sample has smaller mean square error than the sampling distribution of median of a sample.


 Question 10.2.

```{r}
pop2 <- read.csv("C:/Users/ds3/Desktop/yhm-coursework/week2/pop2.csv")
ex2 <- read.csv("C:/Users/ds3/Desktop/yhm-coursework/week2/ex2.csv")

head(pop2)
head(ex2)

n_ex2 <- nrow(ex2)

ex2 |>
        select(group) |>
        filter(group == "HIGH") |>
        summarize(count_high = n())

ex2_prop_high_pressure <- 37/150
ex2_prop_high_pressure   # 0.2466667

pop2 |>
        select(group) |>
        filter(group == "HIGH") |>
        summarize(count_high = n())

pop2_prop_high_pressure <- 28126/ nrow(pop2)
pop2_prop_high_pressure  # 0.28126


P_hat <- rep(0, 10^5)
for (i in 1: 10^5) {
        X <- sample(pop2$group, 150)
        P_hat[i] <- mean(X == "HIGH")
}

mean(P_hat) # 0.2812811
var(P_hat)  # 0.001344962


p <- mean(pop2$group == "HIGH")
p*(1-p)/150
```