---
title: "Predict Citibike"
date: '`r Sys.time()`'
author : "Ye Htut Maung"
output:
  html_document:
    #code_folding: hide
    number_sections: yes
    toc: yes
    toc_depth: 3

---



## Libraries
```{r}
library(tidyverse)
library(modelr)
```

## Loading Data
```{r}
trips_per_day <- read.csv("./week3/trips_per_day.tsv", sep = "\t")
head(trips_per_day)
```

## weekday and weekend
```{r}

trips_per_day<- trips_per_day |>
    mutate(is_weekend = as.numeric(wday(ymd) == 1 | wday(ymd) ==7 ))

```

## Season
```{r}
trips_per_day <- trips_per_day |>
    mutate(season = case_when(
        month(ymd) >= 3 & month(ymd) <= 5 ~ 1,
        month(ymd) >= 6 & month(ymd) <= 8 ~ 2,
        month(ymd) >= 9 & month(ymd) <= 11 ~ 3,
        month(ymd) == 2 ~ 4,
        month(ymd) == 1 ~ 4,
        month(ymd) == 12 ~ 4,
    )) |>
    mutate(season = as.factor(season))
```


## Spliting in Training, and Testing Data
```{r}
set.seed(42)
num_days <- nrow(trips_per_day)
frac_train_test <- 0.9
num_train_test <- floor(num_days * frac_train_test)

# randomly sample rows for the training set
ndx <- sample(1 : num_days, num_train_test, replace = F)

trips_per_day_train <- trips_per_day[ndx, ]
trips_per_day_test <- trips_per_day[-ndx, ]

# K-fold
num_folds <- 5

trips_per_day_train <- trips_per_day_train |>
    mutate(fold = (row_number() %% num_folds) + 1)


```


## Minimum Tempurature and Number of Trips model
```{r}

K <- 1:10

avg_validate_err <- c()
se_validate_err <- c()


for(k in K) {

    validate_err <- c()
    for (f in 1:num_folds) {
        # fit on the training data
        model <- lm(num_trips ~ poly(tmin, k, raw=T), data=filter(trips_per_day_train, fold != f))

        # evaluate on the validation data
        trips_per_day_validate <- filter(trips_per_day_train, fold == f)
        validate_err[f] <- sqrt(mean((predict(model, trips_per_day_validate) - trips_per_day_validate$num_trips)^2))
    }
  avg_validate_err[k] <- mean(validate_err)
  se_validate_err[k] <- sd(validate_err) / sqrt(num_folds)
}

avg_validate_err
se_validate_err
```


## Plot the validate error for minimum Tempurature and Number of Trips model
```{r}
# plot the validate error, highlighting the value of k with the lowest average error
plot_data <- data.frame(K, avg_validate_err, se_validate_err)
ggplot(plot_data, aes(x=K, y=avg_validate_err)) +
  geom_pointrange(aes(ymin=avg_validate_err - se_validate_err,
                      ymax=avg_validate_err + se_validate_err,
                      color=avg_validate_err == min(avg_validate_err))) +
  geom_line(color = "red") +
  scale_x_continuous(breaks=1:12) +
  theme(legend.position="none") +
  xlab('Polynomial Degree') +
  ylab('RMSE on validation data')
```

polynomial for tmin is 4


## Prcp and Number of Trips model


```{r}

K <- 1:10

avg_validate_err_prcp <- c()
se_validate_err_prcp <- c()


for(k in K) {

    validate_err <- c()
    for (f in 1:num_folds) {
        # fit on the training data
        model <- lm(num_trips ~ poly(prcp, k, raw=T), data=filter(trips_per_day_train, fold != f))

        # evaluate on the validation data
        trips_per_day_validate <- filter(trips_per_day_train, fold == f)
        validate_err[f] <- sqrt(mean((predict(model, trips_per_day_validate) - trips_per_day_validate$num_trips)^2))
    }
  avg_validate_err_prcp[k] <- mean(validate_err)
  se_validate_err_prcp[k] <- sd(validate_err) / sqrt(num_folds)
}

avg_validate_err_prcp
se_validate_err_prcp

```


## Plot the validate error for prcp and Number of Trips model
```{r}
# plot the validate error, highlighting the value of k with the lowest average error
plot_data <- data.frame(K, avg_validate_err_prcp, se_validate_err_prcp)
ggplot(plot_data, aes(x=K, y=avg_validate_err_prcp)) +
  geom_pointrange(aes(ymin=avg_validate_err_prcp - se_validate_err_prcp,
                      ymax=avg_validate_err_prcp + se_validate_err_prcp,
                      color=avg_validate_err_prcp == min(avg_validate_err_prcp))) +
  geom_line(color = "red") +
  scale_x_continuous(breaks=1:12) +
  theme(legend.position="none") +
  xlab('Polynomial Degree') +
  ylab('RMSE on validation data')
```

polynomial for prcp is 2




## snwd and Number of Trips model

```{r}

K <- 1:10

avg_validate_err_snwd <- c()
se_validate_err_snwd <- c()


for(k in K) {

    validate_err <- c()
    for (f in 1:num_folds) {
        # fit on the training data
        model <- lm(num_trips ~ poly(snwd, k, raw=T), data=filter(trips_per_day_train, fold != f))

        # evaluate on the validation data
        trips_per_day_validate <- filter(trips_per_day_train, fold == f)
        validate_err[f] <- sqrt(mean((predict(model, trips_per_day_validate) - trips_per_day_validate$num_trips)^2))
    }
  avg_validate_err_snwd[k] <- mean(validate_err)
  se_validate_err_snwd[k] <- sd(validate_err) / sqrt(num_folds)
}

avg_validate_err_snwd
se_validate_err_snwd

```


## Plot the validate error for snwd and Number of Trips model
```{r}
# plot the validate error, highlighting the value of k with the lowest average error
plot_data <- data.frame(K, avg_validate_err_snwd, se_validate_err_snwd)
ggplot(plot_data, aes(x=K, y=avg_validate_err_snwd)) +
  geom_pointrange(aes(ymin=avg_validate_err_snwd - se_validate_err_snwd,
                      ymax=avg_validate_err_snwd + se_validate_err_snwd,
                      color=avg_validate_err_snwd == min(avg_validate_err_snwd))) +
  geom_line(color = "red") +
  scale_x_continuous(breaks=1:12) +
  theme(legend.position="none") +
  xlab('Polynomial Degree') +
  ylab('RMSE on validation data')
```

polynomial for snwd is 5


## snow and Number of Trips model

```{r}

K <- 1:10

avg_validate_err_snow <- c()
se_validate_err_snow <- c()


for(k in K) {

    validate_err <- c()
    for (f in 1:num_folds) {
        # fit on the training data
        model <- lm(num_trips ~ poly(snow, k, raw=T), data=filter(trips_per_day_train, fold != f))

        # evaluate on the validation data
        trips_per_day_validate <- filter(trips_per_day_train, fold == f)
        validate_err[f] <- sqrt(mean((predict(model, trips_per_day_validate) - trips_per_day_validate$num_trips)^2))
    }
  avg_validate_err_snow[k] <- mean(validate_err)
  se_validate_err_snow[k] <- sd(validate_err) / sqrt(num_folds)
}

avg_validate_err_snow
se_validate_err_snow

```


## Plot the validate error for snow and Number of Trips model
```{r}
# plot the validate error, highlighting the value of k with the lowest average error
plot_data <- data.frame(K, avg_validate_err_snow, se_validate_err_snow)
ggplot(plot_data, aes(x=K, y=avg_validate_err_snow)) +
  geom_pointrange(aes(ymin=avg_validate_err_snow - se_validate_err_snow,
                      ymax=avg_validate_err_snow + se_validate_err_snow,
                      color=avg_validate_err_snow == min(avg_validate_err_snow))) +
  geom_line(color = "red") +
  scale_x_continuous(breaks=1:12) +
  theme(legend.position="none") +
  xlab('Polynomial Degree') +
  ylab('RMSE on validation data')
```

polynomial for snow is 2 but we need to think if we should include because its rmse is high



## tmax and Number of Trips model

```{r}

K <- 1:10

avg_validate_err_tmax <- c()
se_validate_err_tmax <- c()


for(k in K) {

    validate_err <- c()
    for (f in 1:num_folds) {
        # fit on the training data
        model <- lm(num_trips ~ poly(tmax, k, raw=T), data=filter(trips_per_day_train, fold != f))

        # evaluate on the validation data
        trips_per_day_validate <- filter(trips_per_day_train, fold == f)
        validate_err[f] <- sqrt(mean((predict(model, trips_per_day_validate) - trips_per_day_validate$num_trips)^2))
    }
  avg_validate_err_tmax[k] <- mean(validate_err)
  se_validate_err_tmax[k] <- sd(validate_err) / sqrt(num_folds)
}

avg_validate_err_tmax
se_validate_err_tmax

```


## Plot the validate error for tmax and Number of Trips model
```{r}
# plot the validate error, highlighting the value of k with the lowest average error
plot_data <- data.frame(K, avg_validate_err_tmax, se_validate_err_tmax)
ggplot(plot_data, aes(x=K, y=avg_validate_err_tmax)) +
  geom_pointrange(aes(ymin=avg_validate_err_tmax - se_validate_err_tmax,
                      ymax=avg_validate_err_tmax + se_validate_err_tmax,
                      color=avg_validate_err_tmax == min(avg_validate_err_tmax))) +
  geom_line(color = "red") +
  scale_x_continuous(breaks=1:12) +
  theme(legend.position="none") +
  xlab('Polynomial Degree') +
  ylab('RMSE on validation data')
```

polynomial for tmax is 8





# Model Selection
```{r}


model_formulas <- c(
    "prcp + snwd + snow + tmax + tmin",
    "poly(prcp, 2, raw=T) + poly(snwd, 2, raw=T) + poly(snow, 2, raw=T) + poly(tmax, 2, raw=T) + poly(tmin, 4, raw=T) ",
    "poly(prcp, 2, raw=T) + poly(snwd, 5, raw=T) + poly(snow, 2, raw=T) + poly(tmax, 8, raw=T) + poly(tmin, 4, raw=T) ",
    "poly(prcp, 2, raw=T) + snwd + poly(tmax, 8, raw=T) + poly(tmin, 4, raw=T)",
    "poly(prcp, 2, raw=T) + poly(tmax, 8, raw=T) + snwd * poly(tmin, 4, raw=T)",
    "poly(prcp, 2, raw=T) + poly(snwd, 5, raw=T) + poly(tmax, 8, raw=T) + poly(tmin, 4, raw=T)",
    "poly(prcp, 2, raw=T) + poly(snwd, 5, raw=T) + poly(tmax, 8, raw=T) + log(tmin)",
    "poly(prcp, 2, raw=T) + sqrt(snwd) + poly(tmax, 8, raw=T) + log(tmin)+ poly(snwd, 5, raw=T)",
    "poly(prcp, 2, raw=T) + sqrt(snwd) + poly(tmax, 8, raw=T) + poly(tmin, 4, raw=T)",
    "prcp + snwd + snow + tmax + tmin + poly(prcp, 2, raw=T) + poly(snwd, 5, raw=T) + poly(tmax, 8, raw=T) + poly(tmin, 4, raw=T)",
    "poly(prcp, 2, raw=T) + poly(snwd, 5, raw=T) + poly(tmax, 8, raw=T) + poly(tmin, 4, raw=T) + is_weekend",
    "poly(prcp, 2, raw=T) + poly(snwd, 5, raw=T) + poly(tmax, 8, raw=T) + poly(tmin, 4, raw=T) + is_weekend + season",
    "poly(prcp, 2, raw=T) + poly(snwd, 5, raw=T) + poly(tmax, 8, raw=T) + poly(tmin, 4, raw=T) + is_weekend * season",
    "poly(prcp, 2, raw=T) + poly(snwd, 5, raw=T) + snow + poly(tmax, 8, raw=T) + poly(tmin, 4, raw=T) + is_weekend * season ",
        "poly(prcp, 2, raw=T) + poly(snwd, 5, raw=T) + poly(tmax, 8, raw=T) + poly(tmin, 4, raw=T) + is_weekend * season "
)

model_formulas = paste0("num_trips ~ ", model_formulas)
M = length(model_formulas)


avg_train_err_1 <- c()
se_train_err_1 <- c()

avg_validate_err_1 <- c()
se_validate_err_1 <- c()



for (m in 1:M) {

    train_err <- c()
    validate_err <- c()
    for (f in 1:num_folds) {
        # fit on the training data
        model <- lm(model_formulas[m], data=filter(trips_per_day_train, fold != f))
        train_err[f] <- sqrt(mean((predict(model, trips_per_day_train) - trips_per_day_train$num_trips)^2))

        # evaluate on the validation data
        trips_per_day_validate <- filter(trips_per_day_train, fold == f)
        validate_err[f] <- sqrt(mean((predict(model, trips_per_day_validate) - trips_per_day_validate$num_trips)^2))
    }

    avg_train_err_1[m] <- mean(train_err)
    se_train_err_1[m] <- sd(train_err) / sqrt(num_folds)

    avg_validate_err_1[m] <- mean(validate_err)
    se_validate_err_1[m] <- sd(validate_err) / sqrt(num_folds)
        
}


# avg_validate_err_1
# se_validate_err_1

for(m in 1:M) {
    print(model_formulas[m])
    print("train error")
    print(avg_train_err_1[m])

    print("validate error")
    print(avg_validate_err_1[m])
    print("------------------------------")
}
```

## Best model fit
best model = "num_trips ~ poly(prcp, 2, raw=T) + poly(snwd, 5, raw=T) + snow + poly(tmax, 8, raw=T) + poly(tmin, 4, raw=T) + is_weekend * season "
```{r}
best_mod = lm(num_trips ~ poly(prcp, 2, raw=T) + poly(snwd, 5, raw=T) + snow + poly(tmax, 8, raw=T) + poly(tmin, 4, raw=T) + is_weekend * season, data = trips_per_day_train)
best_mod
rmse <- sqrt(mean((predict(model, trips_per_day_train) - trips_per_day_train$num_trips)^2))
rmse
rsq <- summary(best_mod)$r.squared
rsq
```



## Plot the final best fit model

```{r}
trips_predictions <- trips_per_day_train |>
    add_predictions(model) |>
    mutate(pred = round(pred))

trips_predictions |>
    ggplot(aes(x = date)) +
    geom_point(aes(y = num_trips, color = "actual trip")) +
    geom_line(aes(y = pred, color= "predicted trip"))
```

```{r}
trips_predictions |>
    ggplot() +
    geom_point(aes(x = pred, y= num_trips)) +
    geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed")

```

```{r}
save(best_mod, file="model.Rdata")
```